Microxanox - an R package for simulating an $MICR$obial ecosystem that can occupy $OX$ic or $ANOX$ic states.

### 2022-04-11 inst/manuscript/manuscript_microxanox.Rmd
Use of elsevier article class and beginning of adaptation to SoftwareX Article Structure

---
title: "Microxanox - an R package for simulating an $MICR$obial ecosystem that can occupy $OX$ic or $ANOX$ic states."
author:
- name: Rainer M Krug
  email: Rainer.Krug@uzh.ch & Rainer@krugs.de
  affiliation: University of Zürich
  footnote: Corresponding Author
- name: Owen L. Petchey
  email: Owen.Petchey@ieu.uzh.ch
  affiliation: University of Zürich
address:
- code: University of Zürich
  address: Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse
    190, 8057 Zurich
keywords: metadata quality; data curation; archival; long term storage; R package;
abstract: | 
  1. set the context for and purpose of the work;
  2. indicate the approach and methods;
  3. outline the main results;
  4. identify the conclusions and the wider implications.
bibliography: manuscript_bibliography.bib

date: last-modified 
date-format: "DD MMM YYYY, HH:mm"

# RMarkdown output settings
output:
  bookdown::html_document2:
    toc: true 
    toc_float:
      collapsed: true
      smooth_scroll: true
  bookdown::pdf_document2:
    keep_tex: yes

# qarto output settings
format:
  html:
    self-contained: true
    toc: true
    toc-depth: 3
    link-external-icon: true
    title-block-style: default
  pdf:
    toc: true
    toc-depth: 3
    lof: true
    include-in-header:
      - headers.tex

# Citation settings    
nocite: '@Bush2017'
linkcolor: blue
link-citations: yes
csl: methods-in-ecology-and-evolution.csl
---

@@

# Introduction

{>>Biological justification and background - one paragraph<<}   
{>>Based on abstract from Romana<<}

Many ecosystems are exposed to gradual changes of environmental variables, to which the responses are not always as gradual as the change of the environmental variable. One example where the gradual change of a single environmental causing an abrupt change of the system is the switch from an aerobic to anaerobic system. This system hass been investigated by @Bush2017. We wanted to take this investigation one step further, and look at the role of an increased biodiversity plays in these dynamics [@REF_NEEDED]. For this purpose, we developed this package.

@@

@@

{>> Add a more detailed description of the model and the additions <<}
{>>TODO<<}

@@

The package will be updated based on further research activities. It will be tested continuously for the last recent, current and devel version of R. 
{>>If we want that, we have to include some, even rudimentary, GITHUB tests.<<}
{>>RMK: Agreed - let's discuss this when we talk about the package paper.<<}

@@

```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap='Plot of results of a simulation run using the function `plot_dynamics()`$`. Details can be found in the "User Guide" section "Three strains per functional group".', fig.align="center", out.width = 500 }
knitr::include_graphics("figures/fig1_plot_dynamics.png")
```

@@

The first two use cases come from the User Guide and the Partial Reproduction Vignettes. The third comes from a research article that relied on the package. All of these use cases can be expanded to larger numbers of strains per functional group.

@@

In the [User Guide](LINK_NEEDED) we ued a one strain system (section "1 strain per functional group") and three strain system (section "3 strains per functional group") to analyse as an example the determination of the stable states during temporal environmental chan ge (the oxygen diffusivity). From these simulatiuons, we extracted measures of nonlinearity and hysteresis.

@@

One of the reasons to develop this package was to reproduce the results presented in @Bush2017, this was achieved as demonstrated in the [Partial Reproduction supplement](LINK_NEEDED). All aspects in the paper could be reproduced.

@@

{>>More details from the paper.<<}  
{>>Romana will provide two or three sentences<<}

>>>>

---
title: "Microxanox - an R package for simulating an $MICR$obial ecosystem that can occupy $OX$ic or $ANOX$ic states."
author:
- name: Rainer M Krug
  email: Rainer.Krug@uzh.ch & Rainer@krugs.de
  affiliation: University of Zürich
  footnote: Corresponding Author
- name: Owen L. Petchey
  email: Owen.Petchey@ieu.uzh.ch
  affiliation: University of Zürich
address:
- code: University of Zürich
  address: Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse
    190, 8057 Zurich
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"  
keywords: metadata quality; data curation; archival; long term storage; R package;
abstract: | 
  Ca. 100 words.
bibliography: manuscript_bibliography.bib
journal: "SoftwareX"
date: "`r Sys.Date()`"

#linenumbers: true
#numbersections: true
csl: elsevier-harvard.csl
output: 
  rticles::elsevier_article:
    includes: 
      in_header: headers.tex

# Citation settings    
nocite: '@Bush2017'
linkcolor: blue
link-citations: yes
---

@@

# Required Metadata
## Current code version
Ancillary data table required for subversion of the codebase.

| **Nr.** | **Code metadata description** | **Please fill in this column** |
|---|---|---|
| C1 | Current code version | v0.9.0 |
| C2 | Permanent link to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) |
| C3 | Code Ocean compute capsule |  |
| C4 | Legal Code License   | CC BY 4.0 |
| C5 | Code versioning system used |git |
| C6 | Software code languages, tools, and services used | R | 
| C7 | Compilation requirements, operating environments | R (>= 4.1.0), ADD PACKAGES | 
| C8 | If available Link to developer documentation/manual |  |
| C9 | Support email for questions | Rainer.Krug@uzh.ch; Rainer@krugs.de |

# Motivation and significance
Many ecosystems are exposed to gradual changes of environmental variables, to which the responses are not always as gradual as the change of the environmental variable. One example where the gradual change of a single environmental causing an abrupt change of the system is the switch from an aerobic to anaerobic system. This system hass been investigated by @Bush2017. We wanted to take this investigation one step further, and look at the role of an increased biodiversity plays in these dynamics [@REF_NEEDED]. For this purpose, we developed this package.

@@

# Software description
Describe the software in as much as is necessary to establish a vocabulary needed to explain its impact.

## Software Architecture
Give a short overview of the overall software architecture; provide a pic- torial component overview or similar (if possible). If necessary provide im- plementation details.

## Software Functionalities
Present the major functionalities of the software.

# Illustrative Examples
Provide at least one illustrative example to demonstrate the major functions.
Optional: you may include one explanatory video that will appear next to your article, in the right hand side panel.

# Impact
This is the main section of the article and the reviewers weight the description here appropriately
Indicate in what way new research questions can be pursued as a result of the software (if any).
Indicate in what way, and to what extent, the pursuit of existing research questions is improved (if so).
Indicate in what way the software has changed the daily practice of its users (if so).
Indicate how widespread the use of the software is within and outside the intended user group.
Indicate in what way the software is used in commercial settings and/or how it led to the creation of spin-off companies (if so).

# Conclusions
Set out the conclusion of this original software publication.

# Conflict of Interest
We wish to confirm that there are no known conflicts of interest associated with this publication and there has been no significant financial support for this work that could have influenced its outcome.

# Acknowledgements
Optionally thank people and institutes you need to acknowledge.

# THIS NEEDS TO BE TRANSFERED IN THE FORMAT ABOVE

@@

<span style="color:red">{>> Add a more detailed description of the model and the additions <<}
{>>TODO<<}</span>

@@

The package will be updated based on further research activities. It will be tested continuously for the last recent, current and devel version of R. 
<span style="color:green">{>>If we want that, we have to include some, even rudimentary, GITHUB tests.<<}
{>>RMK: Agreed - let's discuss this when we talk about the package paper.<<}
</span>

@@

```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap='Plot of results of a simulation run using the function $plot_dynamics()$. Details can be found in the "User Guide" section "Three strains per functional group".', fig.align="center", out.width = 500 }
knitr::include_graphics("figures/ug_three_strains_dynamics.png")
```

@@

The first two use cases are described in detail in the User Guide and the Partial Reproduction Vignettes. The third is taken from the @REF_NEEDED for which this R package was designed. All of these use cases can be expanded to larger numbers of strains per functional group and variable values van be changed.

@@

In the [User Guide](LINK_NEEDED) we used a one strain system (section "1 strain per functional group") and three strain system (section "3 strains per functional group") to determine as an example the stable states during temporal environmental changes (the oxygen diffusivity). From these simulatiuons, we extracted measures of nonlinearity and hysteresis. See Fig \@ref(fig:plot-dynamics) as an example plot of the simulations.

```{r uc1_stable_state, echo = FALSE,  include=TRUE, fig.cap='Plot of the stable states of the simulation runs under different oxygen diffusivity. The top graph are the Organisms (ech initially with three strains) while the lower graph is the substrate availability under the same oxygen diffusivities. Details can be found in the "User Guide" section "Three strains per functional group".', fig.align="center", out.width = 500 }
knitr::include_graphics("figures/ug_three_strains_stable_state.png")
```

@@

One of the reasons to develop this package was to reproduce the results presented in @Bush2017, this was achieved as demonstrated in the [Partial Reproduction supplement](LINK_NEEDED). All aspects in the paper could be reproduced and are shown in the vignette.

```{r user_guide_hysteresis, echo = FALSE,  include=TRUE, fig.cap='Hysteresis of all assessed combinations of variability.', fig.align="center", out.width = 500 }
knitr::include_graphics("figures/user_guide_hysteresis.png")
```

@@

<span style="color:green">{>>More details from the paper.<<}  
{>>Romana will provide two or three sentences<<}</span>
